\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-10 k-planar.tex
Main theorem in appendix is now proved, subject to  a tweak to the main lemma

\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\depth}{depth}
\DeclareMathOperator{\ff}{f}
\DeclareMathOperator{\tw}{tw}
\DeclareMathOperator{\pw}{pw}
\DeclareMathOperator{\qn}{qn}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

@@

\newcommand{\PP}{\mathcal{P}}
\renewcommand{\SS}{\mathcal{S}}

@@

A \defin{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle L_0,L_1,\ldots\rangle$ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$ and for every edge $vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.  For any partition $\PP=\{S_1,\ldots,S_p\}$ of $V(G)$, a \defin{quotient graph} $H=G/\PP$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$. To highlight the importance of the quotient graph $H$, we call $\PP$ an \defin{$H$-partition} and write this concisely as $\PP=\{S_x : x\in V(H)\}$ so that each element of $\PP$ is indexed by the vertex it creates in $H$.

@@

\note{DW}{Work in progress}

@@

For a cycle $C$, we write $C = [ P_1 , \dots , P_k ]$ if $P_1 , \dots, P_k$ are pairwise disjoint non-empty paths in $C$, and the endpoints of each path $P_i$ can be labelled $x_i$ and $y_i$ so that $y_ix_{i+1} \in E(C)$ for $i\in\{1,\dots,k\}$, where $x_{k+1}$ means
$x_1$. This implies that $\{P_1,\dots,P_k\}$ is a vertex-partition of $C$. \note{DW}{Allow $P_i$ to be empty?}

If $Q$ is a path in a graph, and $P_1, \dots,P_k$ are edge-disjoint non-empty subpaths of $Q$ ordered by $Q$ and partitioning $E(Q)$, then we write $Q=P_1,\dots,P_k$.

\note{DW}{It would help to say that we consider a path to be an (unordered) subgraph. Would that be a problem anywhere else?}

@@

Say $G$ is a 2-connected plane multigraph. Let $T$ be a spanning tree of $G$ rooted at a vertex $r$ on the outerface of $G$. A \defin{bipod} in $G$ (with respect to $T$) is any path $P_1SP_2$, where $S$ is a facial path in $G$, and each $P_i$ is a vertical path in $T$ whose lower endpoint is in $S$. Here it is possible that $S$ or $P_i$ is empty. So every subpath of a bipod is a bipod. A \defin{tripod} in $G$ (with respect to $T$) is either a bipod in $G$ or a subgraph of $G$ consisting of a facial cycle $F$ of $G$, plus three distinct vertical paths $P_1,P_2,P_3$ in $T$ where the lower endpoint of each $P_i$ is in $F$. \note{DW}{Need to think about degenerate cases.}. These definitions generalise definitions in \citep{DJMMUW20} where $G$ is assumed to be a plane triangulation, in which case $F$ is a 3-cycle. \note{DW}{We actually managed to omit the definition of `bipod' in \citep{DJMMUW20}!}

@@

\begin{lem}
Let $G_0$ be a 2-connected plane multigraph. Let $T$ be a spanning tree of $G_0$ rooted at some vertex $r$ on the boundary of the outerface of $G_0$. Let $k\in\{1,2,3\}$ and let $C = [P_1,\dots,P_k]$ be a cycle in $G_0$ with $r$ in the exterior of $C$, where $P_1, \dots, P_k$ are bipods in $G_0$. 
	\note{DW}{We need to assume that $|V(C)|\geq 3$ and ensure that when we apply induction $|V(Z_j)|\geq 3$. I think it is okay, but it needs an argument.}
	Let $G$ be the plane graph consisting of all the edges and vertices of $G_0$ in $C$ and the interior of $C$. Then $G$ has a partition $\PP$ into tripods with respect to $T$ such that:
\begin{compactitem}
\item $P_1,\dots,P_k\in\PP$,
\item $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that:
\begin{compactitem}
	\item some bag of $\mathcal{T}$ contains $P_1,\dots,P_k$, and 
	\item for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
\end{compactitem}
\end{compactitem}
\end{lem}

\begin{proof}
\note{DW}{Can we assume $k=3$ by allowing $P_i=\emptyset$? That would clean up the proof. We could scrap $R_1,R_2,R_3$ and just use $P_1,P_2,P_3$.}

The proof is by induction on $|V(G)\setminus V(C)|$. If $V(G)\setminus V(C)=\emptyset$, then $G$ has just one face bounded by $C$ and the desired partition is $\PP=\{P_1,\ldots,P_k\}$.  The desired tree-decomposition of $G/\PP$ consists of a single bag $\{P_1,\ldots,P_k\}$ indexed by a 1-node tree.

Now assume that $V(G)\setminus V(C)\neq\emptyset$. Partition the vertices of $C$ into paths $R_1$, $R_2$, and $R_3$ as follows:
\begin{compactenum}
	\item If $k=1$ then, since $C$ is a cycle, $P_1$ has at least three vertices, so $P_1=[v, P_1', w]$ for two distinct vertices $v$ and $w$. We set $R_1:=v$, $R_2:=P_1'$ and $R_3:=w$.

@@

@@

Let $Y$ be the tripod $F\cup Q_1'\cup Q_2' \cup Q_3'$. Say $F=v_1 S_1 v_2 S_2 v_3 S_3$, where $S_i$ is a facial path in $F$ clockwise from $v_i$ to $v_{i+1}$ (where $v_4=v_1$). Let $L_i$ be the path $Q_i'S_iQ_{i+1}'$ for $i\in\{1,2,3\}$ (where $Q_4=Q_1$ and $R_4=R_1$). So $L_i$ is a bipod. \note{DW}{Note that $S_i\neq\emptyset$ since $v_i\neq v_{i+1}$.}

@@

\note{DW}{For a cycle $C$ in a plane graph $G$, define $G\langle C\rangle$ to be the subgraph of $G$ consisting of $C$ and all the edges and vertices of $G$ in the interior of $C$?}

Consider $j\in\{1,\dots,m\}$. Let $G_j$ be the subgraph of $G$ consisting of $Z_j$ and all the edges and vertices of $G$ in the interior of $Z_j$. The outerface of $G_j$ is $Z_j$. Observe that for some $i\in\{1,2,3\}$, we have
$Z_j=[R_{i,j}^+,R_{i+1,j}^-,S_{i,j}]$, where 
$R_{i,j}^+$ is some subpath of $R_i^+$,
$R_{i+1,j}^-$ is some subpath of $R_{i+1}^-$, and
$L_{i,j}$ is the path $Z_j-V(C)$ (which is contained in $Y$). 
Since $R_i^+$, $R_{i+1}^-$ and $L_i$ are bipods, 
$R_{i,j}^+$, $R_{i+1,j}^-$ and $L_{i,j}$ are bipods. 
By construction, $v_{i-1}$ is not in $G_j$ (where $v_0=v_3$). 
Thus $V(G_j)\setminus V(Z_j) \subsetneq V(G)\setminus V(C)$, and we may apply induction to $G_j$.  Thus $G_j$ has a partition $\PP_j$ into tripods with respect to $T$ such that:
\begin{compactitem}
	\item $R_{i,j}^+, R_{i+1,j}^-, L_{i,j}\in\PP_j$, 
	\item the quotient $G_j / \PP_j$ has a tree-decomposition $\mathcal{T}_j=\{B_x:x\in V(D_j)\}$ of width at most 3 such that: 
	\begin{compactitem}
		\item some bag $B_{u_j}$ of $\mathcal{T}_j$ contains $R_{i,j}^+, R_{i+1,j}^-, L_{i,j}$, and 
		\item for every facial cycle $F'$ of $G_j$ there is a bag $B$ of $\mathcal{T}_j$ such that $V(F')\subseteq \bigcup_{X\in B} X$.
	\end{compactitem}
\end{compactitem}

@@

We now construct the desired partition $\PP$ of $G$. 
Let $\PP$ be obtained from $\bigcup_{j\in\{1,\dots,m\}} \PP_j$ by 
replacing $R_{i,j}^+\in\PP_j$ by $P_i$, 
replacing $R_{i+1,j}^-\in\PP_j$ by $P_{i+1}$, and
replacing $L_{i,j}\in\PP_j$ by $Y$. \note{DW}{What if $k\neq 3$? Then $P_2$ or $P_3$ might not be defined.}
By construction, $\PP$ partitions $V(G)$ into tripods with respect to $T$ and $P_1,\ldots,P_k\in \PP$.

@@

Every facial cycle $F_0$ of $G$ is a face of some $G_j$ or is $F$. In the latter case, $V(F)\subseteq Y \subseteq \bigcup_{X\in B_u} X$. In the first case, there is a bag $B_x$ of $\mathcal{T}_j$ such that $V(F)\subseteq \bigcup_{X\in B_x} X$. This property is maintained by the above replacements since 
$R_{i,j}^+ \subseteq P_i$, 
$R_{i+1,j}^- \subseteq P_{i+1}$, and
$L_{i,j}\subseteq Y$. \note{DW}{What if $k\neq 3$?}

@@

@@

@@

I claim that G_2 is contained in H * P * K_{ d * floor(d/2) } where tw(H) <= 3.

With d=4, G_2 includes all edge-maximal 1-planar  multigraphs,
and G_2 is contained in H * P * K_8 where tw(H) <= 3.

Let T be a bfs spanning tree of G_0 rooted on the outerface.

Run the original proof, finding three paths Q_1,Q_2,Q_3
leading to a Sperner triangle t in G_1, which is a subset of a face F of G_0.
The corresponding tripod in our partition is Q_1 \cup Q_2 \cup Q_3 \cup F.
We can apply induction inside the regions bounded by C, Q_1, Q_2, Q_3 and F.

@@

This setup (if it works) also works for map graphs in which every face
is a nation.

I just thought we need to be careful with the vertex 3-colouring.
If v is on the inside of C and the path in T from v to the root
first hits C at some B_i, then what is the colour of v?
This is not a problem there is a noose in G_0 including A_1..A_k, so
the path in T from v to the root must intersect A_1...A_k.

@@

	\begin{figure}[!b]
	\begin{center}
		\begin{tabular}{cc}
			\includegraphics[scale=1.4]{induction-i-3} &
			\includegraphics[scale=1.4]{induction-i-1} \\
			(a) & (b) \\[1em]
			\includegraphics[scale=1.4]{induction-i-2} &
			\includegraphics[scale=1.4]{induction-iii} \\
			(c) & (d) 
		\end{tabular}
	\end{center}
	\caption{The inductive proof of \cref{NearTriang}: (a)~the spanning tree $T$ and the paths $P_1,\ldots,P_4$; 
		(b) the paths $R_1$, $R_2$, $R_3$, and the Sperner triangle $\tau$; 
		(c)~the paths $Q_1'$, $Q_2'$ and $Q_3'$; 
		(d)~the near-triangulations $G_1$, $G_2$, and $G_3$, with the vertical paths of $T$ on $F_1$, $F_2$, and $F_3$.}
	\label{NearTriangProof}
\end{figure}

>>>>

\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\depth}{depth}
\DeclareMathOperator{\ff}{f}
\DeclareMathOperator{\tw}{tw}
\DeclareMathOperator{\pw}{pw}
\DeclareMathOperator{\qn}{qn}

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
\DeclarePairedDelimiter{\blah}{\langle}{\rangle}

@@

\newcommand{\TT}{\mathcal{T}}
\newcommand{\PP}{\mathcal{P}}
\renewcommand{\SS}{\mathcal{S}}

@@

A \defin{layering} of a graph $G$ is a sequence $\mathcal{L}=(L_0,L_1,\ldots)$ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$ and for every edge $vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.  For any partition $\PP=\{S_1,\ldots,S_p\}$ of $V(G)$, a \defin{quotient graph} $H=G/\PP$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$. To highlight the importance of the quotient graph $H$, we call $\PP$ an \defin{$H$-partition} and write this concisely as $\PP=\{S_x : x\in V(H)\}$ so that each element of $\PP$ is indexed by the vertex it creates in $H$.

@@

@@

For a cycle $C$, we write $C = [ P_1 , \dots , P_k ]$ if $P_1 , \dots, P_k$ are pairwise disjoint (possibly empty) path subgraphs in $C$ such that $V(C)=\bigcup_i V(P_i)$ and $P_1,\dots,P_k$ are ordered around $C$.

%and the endpoints of each path $P_i$ can be labelled $x_i$ and $y_i$ so that $y_ix_{i+1} \in E(C)$ for $i\in\{1,\dots,k\}$, where $x_{k+1}$ means $x_1$. This implies that $\{P_1,\dots,P_k\}$ is a vertex-partition of $C$. \note{DW}{In the proof below, when we apply induction, we need to allow $P_i=\emptyset$, right?}

If $Q$ is a path in a graph, and $P_1, \dots,P_k$ are edge-disjoint non-empty subpaths of $Q$ ordered by $Q$ and partitioning $E(Q)$, then we write $Q=P_1,\dots,P_k$. \note{DW}{Do we need this?}

@@

Say $G$ is a 2-connected plane multigraph. Let $T$ be a spanning tree of $G$ rooted at a vertex $r$ on the outerface of $G$. A \defin{bipod} in $G$ (with respect to $T$) is any path $P_1SP_2$ (considered to be a subgraph), where $S$ is a facial path in $G$, and each $P_i$ is a vertical path in $T$ whose lower endpoint is in $S$. Here it is possible that $S$ or $P_i$ is empty. So every subpath of a bipod is a bipod. A \defin{tripod} in $G$ (with respect to $T$) is either a bipod in $G$ or a subgraph of $G$ consisting of a facial cycle $F$ of $G$, plus three distinct vertical paths $P_1,P_2,P_3$ in $T$ where the lower endpoint of each $P_i$ is in $F$. These definitions generalise definitions in \citep{DJMMUW20} where $G$ is assumed to be a plane triangulation, in which case $F$ is a 3-cycle. \note{DW}{We actually managed to omit the definition of `bipod' in \citep{DJMMUW20}!}

@@

For a cycle $C$ in a plane multigraph $G$, define \defin{$G\blah{C}$} to be the subgraph of $G$ consisting of $C$ and all the edges and vertices of $G$ in the interior of $C$.

\begin{lem}
	\label{NewLemma}
Let $G_0$ be a 2-connected plane multigraph. Let $T$ be a spanning tree of $G_0$ rooted at some vertex $r$ on the boundary of the outerface of $G_0$. Let $C = [P_1,\dots,P_k]$ be a cycle in $G_0$ with $r$ in the exterior of $C$, where $k\in\{1,2,3\}$ and $P_1, \dots, P_k$ are bipods in $G_0$. Let $G:=G_0\blah{C}$. Then $G$ has a partition $\PP$ into tripods with respect to $T$ such that:
\begin{compactitem}
\item $P_1,\dots,P_k\in\PP$,
\item $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that:
\begin{compactenum}[(i)]
	\item some bag of $\mathcal{T}$ contains $P_1,\dots,P_k$, and 
	\item for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
\end{compactenum}
\note{DW}{We need to replace (ii) by ``for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$''. Curiously, this implies (i) (since $C$ is a facial cycle of $G$).}
\end{compactitem}
\end{lem}

\begin{proof}
We may assume that each of $P_1,\dots,P_k$ is non-empty (since if some $P_i$ is empty, then we just consider $C=[P_1,\dots,P_{i-1},P_{i+1},\dots,P_k]$).

The proof is by induction on the number of faces of $G$. If $G$ has just one internal face (which is bounded by $C$), then the desired partition is $\PP=\{P_1,\ldots,P_k\}$ and the desired tree-decomposition of $G/\PP$ consists of a single bag $\{P_1,\ldots,P_k\}$ indexed by a 1-node tree. Now assume that $G$ has at least two internal faces.

First suppose that $C$ is a 2-cycle. Thus $k\leq 2$. Let $e$ be one of the edges in $C$. Let $F$ be the internal facial cycle of $G$ that includes $e$. Since each facial cycle of $G_0$ has length at least 3, $F$ consists of a non-empty path $P$ plus the edge $e$. Let $C'$ be obtained by replacing $e$ in $C$ by $P$. Let $G':=G\langle{C'}\rangle$. Thus $G'$ has one fewer face than $G$. Note that $C'=[P_1,\dots,P_k,P]$ (\note{DW}{or it can be written that way?}).  By induction, $G'$ has a partition $\PP$ into tripods with respect to $T$ such that:
\begin{compactitem}
	\item $P_1,\dots,P_k,P\in\PP$,
	\item $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that:
	\begin{compactenum}[(i)]
	\item some bag of $\mathcal{T}$ contains $P_1,\dots,P_k,P$, and 
	\item for every facial cycle $F'$ of $G\blah{C'}$ there is a bag $B$ of $\mathcal{T}$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
	\end{compactenum}
\end{compactitem}
The bag of $\mathcal{T}$ that contains $P_1,\dots,P_k,P$ satisfies property (ii) for $F$. 
Thus, $\PP$ and $\TT$ satisfy the desired properties for $G$. Now assume that $|V(C)|\geq 3$.

%\note{DW}{One option is to do induction on $|V(G)\setminus V(C)|-k$ and reduce to the $k=3$ case (similar to the $|V(C)|\leq 2$ case). When $k<3$ we can grab another face $F$ on the inside of $C$ and make the boundary of $F$ a new path. Then the main proof will only deal with the $k=3$ case. Is this any easier?}

%Now suppose that $k=1$. Let $v$ and $w$ be the first and last vertex in $P_1$. So $C$ consists of $P_1$ plus the edge $vw$. Let $F$ be the internal facial cycle of $G$ containing $vw$.......... Now suppose that $k=2$ .....

Partition the vertices of $C$ into paths $R_1$, $R_2$, and $R_3$ as follows:
\begin{compactenum}
	\item If $k=1$ then, $P_1$ has at least three vertices (since $|V(C)|\geq 3$), so $P_1=[v, P_1', w]$ for two distinct vertices $v$ and $w$. We set $R_1:=v$, $R_2:=P_1'$ and $R_3:=w$.

@@

Note that $R_i \subseteq P_{\alpha(i)}$ for some $\alpha(i)\in \{1,\dots,k\}$.

@@

Let $Y$ be the tripod $F\cup Q_1'\cup Q_2' \cup Q_3'$. Say $F=v_1 S_1 v_2 S_2 v_3 S_3$, where $S_i$ is a facial path in $F$ clockwise from $v_i$ to $v_{i+1}$ (where $v_4=v_1$). Let $L_i$ be the path $Q_i'S_iQ_{i+1}'$ for $i\in\{1,2,3\}$ (where $Q_4=Q_1$ and $R_4=R_1$). So $L_i$ is a bipod. \note{DW}{Note that $S_i\neq\emptyset$ since $v_i\neq v_{i+1}$. Is this useful.}

@@

Consider $j\in\{1,\dots,m\}$. Let $G_j$ be the subgraph of $G$ consisting of $Z_j$ and all the edges and vertices of $G$ in the interior of $Z_j$. The outerface of $G_j$ is $Z_j$. Observe that for some $i\in\{1,2,3\}$, we have
$Z_j=[R_{i,j}^+,R_{i+1,j}^-,S_{i,j}]$, where 
$R_{i,j}^+$ is some subpath of $R_i^+$,
$R_{i+1,j}^-$ is some subpath of $R_{i+1}^-$, and
$L_{i,j}$ is the path $Z_j-V(C)$ (which is contained in $Y$). \note{DW}{Add a figure of a good case and a weird case.}
Since $R_i^+$, $R_{i+1}^-$ and $L_i$ are bipods, 
$R_{i,j}^+$, $R_{i+1,j}^-$ and $L_{i,j}$ are bipods. 
By construction, $v_{i-1}$ is not in $G_j$ (where $v_0=v_3$). 
Thus $V(G_j)\setminus V(Z_j) \subsetneq V(G)\setminus V(C)$, and we may apply induction to $G_j$.  Thus $G_j$ has a partition $\PP_j$ into tripods with respect to $T$ such that:
\begin{compactitem}
	\item $R_{i,j}^+, R_{i+1,j}^-, L_{i,j}\in\PP_j$, 
	\item the quotient $G_j / \PP_j$ has a tree-decomposition $\mathcal{T}_j=\{B_x:x\in V(D_j)\}$ of width at most 3 such that: 
	\begin{compactitem}
		\item some bag $B_{u_j}$ of $\mathcal{T}_j$ contains $R_{i,j}^+, R_{i+1,j}^-, L_{i,j}$, and 
		\item for every facial cycle $F'$ of $G_j$ there is a bag $B$ of $\mathcal{T}_j$ such that $V(F')\subseteq \bigcup_{X\in B} X$.
	\end{compactitem}
\end{compactitem}

@@

We now construct the desired partition $\PP$ of $G$. 
Let $\PP$ be obtained from $\bigcup_{j\in\{1,\dots,m\}} \PP_j$ by 
replacing $R_{i,j}^+\in\PP_j$ by $P_{\alpha(i)}$, 
replacing $R_{i+1,j}^-\in\PP_j$ by $P_{\alpha(i+1)}$, and
replacing $L_{i,j}\in\PP_j$ by $Y$. 
By construction, $\PP$ partitions $V(G)$ into tripods with respect to $T$ and $P_1,\ldots,P_k\in \PP$.

@@

Every facial cycle $F_0$ of $G$ is a face of some $G_j$ or is $F$. In the latter case, $V(F)\subseteq Y \subseteq \bigcup_{X\in B_u} X$. In the first case, there is a bag $B_x$ of $\mathcal{T}_j$ such that $V(F)\subseteq \bigcup_{X\in B_x} X$. This property is maintained by the above replacements since 
$R_{i,j}^+ \subseteq P_{\alpha(i)}$ and 
$R_{i+1,j}^- \subseteq P_{\alpha(i+1)}$ and
$L_{i,j}\subseteq Y$.

@@

\begin{proof}[Proof of \cref{CliqueFaceProductStructure}] 
Let $G$ be a 2-connected plane multigraph, in which every facial cycle has length in $\{3,\dots,d\}$. Let $C$ be the outer-cycle of $G$. Let $G^+$ be obtained from $G$ by adding a new vertex $r$ adjacent to every vertex on the outerface of $G$. So $G^+$ is a 2-connected plane multigraph and $G=G^+\blah{C}$. Let $T$ be a BFS spanning tree of $G^+$ rooted at $r$. Let $\mathcal{L}$ be the corresponding BFS layering. By \cref{NewLemma} \note{DW}{with improved (ii)}, $G$ has a partition $\PP$ into tripods with respect to $T$ such that $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$. Since $G$ is a spanning subgraph of $\widehat{G}$, we may consider $\PP$ to be a partition of $\widehat{G}$, and $G/\PP \subseteq \widehat{G}/\PP$. Consider an edge $XY$ of $\widehat{G}/\PP$. By the definition of $\widehat{G}$, there is a facial cycle $F$ of $G$ such that both $X$ and $Y$ intersect $F$. By assumption, there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$. In particular, $X,Y\in B$. Hence $\TT$ is also a tree-decomposition of $\widehat{G}$. Each tripod in $\PP$ has at most $d$ vertices in each layer in $\mathcal{L}$. For each edge $vw$ of $\widehat{G}$, if $v\in L_i$ and $w\in L_j$ then $|i-j|\leq\floor{\frac{d}{2}}$. Grouping blocks of $\floor{\frac{d}{2}}$ consecutive layers gives a layering of $\widehat{H}$ with layered width $\floor{\frac{d}{2}}d$ with respect to $\PP$. By \cref{PartitionProduct}, $\widehat{G}$ is contained in $H \boxtimes P \boxtimes K_{ d  \floor{d/2} }$ where $H:=\widehat{G}/\PP$, which has treewidth 3.
\end{proof}

We now give two graph classes for which \cref{CliqueFaceProductStructure} is applicable.

@@

\begin{lem}
Every $d$-map graph is a subgraph of $\widehat{G}$ for some 2-connected plane graph $G$ in which every facial cycle has length in $\{3,\dots,d\}$.
\end{lem}

@@

@@

@@
